Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-09-02 doc/papers/cgo2014/paper.tex
CGO: dedicated section for alias and effects analyses

The limitation of expression templates presented in
\S~\ref{sec:expression-templates} is easy to solve at the Python compiler
level, using forward substitution. Typically, forward substitution is
supported by the computation of a dependency graph that itself relies on
read-write effects and alias analysis.

Alias analysis is impossible for Python programs in general, because of late
binding and possible monkey patching---a way to modify the code of dynamic
languages at run-time. In the context of a static compiler for
numerical programs, it can be assumed that binding is done statically and no
monkey patching is done. In that context, it is relatively easy to design an
intra-procedural alias analysis thanks to the absence of pointer arithmetic:
the only language construction that can create aliasing is the assignment, thus
a simple control-flow base alias analysis yields accurate results.

@@

Thanks to these analysis, we are able to say that \texttt{t0} and \texttt{t1}
can be forwarded in the Listing~\ref{lst:rosen}.

>>>>

The limitation of expression templates presented in
\S~\ref{sec:expression-templates} can be solved using forward substitution at 
the Python compiler level. This optimization leverages the 
\textit{def-use} chains~\cite{defuse} as well as interprocedural read-write 
effects and alias analysis presented in the next sections. 
Thanks to these analyses, the forward substitution implementation is quite 
simple: \todo{it is not THAT simple indeed, there is "lazyness" analysis 
involved}....

This optimization is able to forward \texttt{t0} and \texttt{t1} in 
Listing~\ref{lst:rosen}, avoiding a temporary allocation and providing more
arithmetic intensity at the same time.

\subsection{Alias Analysis}
\label{sec:alias}

Alias analysis is impossible for Python programs in general, because of late
binding and possible monkey patching---a way to modify the code of dynamic
languages at run-time. In the context of a static compiler for
numerical programs, it can be assumed that binding is done statically and no
monkey patching is done. Indeed the Pythran compiler operates on a subset of 
Python that forbid this kind of dynamic patching. 
In this context, it becomes possible to design an intra-procedural alias 
analysis thanks to the absence of pointer arithmetic:
the only language construction that can create aliasing is the assignment, thus
a simple control-flow based alias analysis yields accurate results.

\subsection{Interprocedural Effects Analysis}
\label{sec:effects}

@@
